---
title: "R for Biodiversity Science"
execute:
  echo: true
format: html
---

# data management

## Read in data in one line

```{r, message=FALSE}
library(tidyverse)
soy <- read_csv("soybean.csv")
soy
```

## Filter

```{r}
soy %>% 
  filter(gen == c("G01", "G02"))
```

## Sort

```{r}
soy %>% 
  arrange(yield)
```

## Select

```{r}
soy %>% 
  select(gen, yield)
```

## Advanced Select

```{r}
soy %>% 
  select(where(is.numeric))
```

## Summarizing data

```{r}
soy %>% 
  group_by(loc, year) %>% 
  summarize(mean_yield = mean(yield), 
            sd_yield = sd(yield))
```

# data visualization

## basic plotting

```{r}
soy %>% 
  ggplot(aes(x = oil, y = protein)) +
  geom_point()
```

## customize your plot

```{r}
soy %>% 
  ggplot(aes(x = oil, y = protein, color = loc)) +
  geom_point(alpha = 0.5)
```

## customize even furthur

```{r}
my_graph <- soy %>% 
  ggplot(aes(x = oil, y = protein, color = loc)) +
  geom_point(alpha = 0.5, size = 3) +
  scale_color_brewer(palette = "Set1") +
  theme_bw(base_size = 15, base_family = "serif") 

my_graph
```

## Create complex graphs

```{r}
my_graph +
  facet_grid(year ~ loc) +
  geom_smooth(color = "blue")
```

## 3D mapping

```{r}
library(plotly)

plot_ly(soy, x = ~protein, y = ~oil, z = ~height, color = ~ factor(year)) %>% 
  add_markers()
```

# Statistics

## comparing groups visually

```{r}
library(ggpubr)

ggbarplot(soy, x = "loc", y = "oil", add = "mean_sd", fill = "loc")
```

## statistical test with ANOVA

```{r}
loc_anova <- aov(oil ~ loc, data = soy)
summary(loc_anova)
```

## post-comparison

```{r}
library(agricolae)
loc_hsd <- HSD.test(loc_anova, trt = "loc")
loc_hsd
```

## plot the posthoc results

```{r}
plot(loc_hsd)
```

# map making

## Download boundary data from online and plot

```{r}
thailand <- map_data("world", "thailand")
ggplot(thailand, aes(long, lat, group = group)) +
  geom_polygon(fill = "brown") +
  coord_quickmap()
```

## plot distribution of our plant

```{r}
maerua <- read_csv("maerua_coord.csv")
ggplot() +
  geom_polygon(aes(long, lat, group = group), fill = "#B0C59D", data = thailand) +
  geom_point(aes(lon, lat), color = "red", data = maerua) +
  coord_quickmap()
```

# diversity / community analysis

## work with community data

```{r, echo=FALSE}
library(vegan)
data(dune)
data(dune.env)
```

```{r}
dune
```

## visualize community data

```{r}
heatmap(as.matrix(dune))
```

## have we got all the species yet?

```{r}
plot(specaccum(dune))
```

## calculate diversity of each plot

Species Number

```{r}
specnumber(dune)
```

Shannon's Index

```{r}
diversity(dune, index = "shannon")
```

## community analysis with PCA

```{r}
library(ggfortify)
dune_pca <- prcomp(dune)
autoplot(dune_pca, colour = "Management", data = dune.env, frame = TRUE)
```
